---
title: "EDA"
output:
  word_document: default
  pdf_document: default
date: "2024-12-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading in all files/packages

```{r}
data <-read.csv("/Users/ahzenthu/Downloads/Bank Customer Churn Prediction.csv")
head(data)
#bc churn is the target variable, lets split the data into churn and not churn
churned <- data[data$churn == 1,]
notchurned <- data[data$churn == 0,]
head(churned)
head(notchurned)
```

## stat summary

```{r}
# for the whole data, broad data check
summary(data)
str(data)
#any missing values?
sum(is.na(data))
anyNA(data)
```
In the broad statistical summary of the dataset, (skipping over id) we see that the data has numerical and categorical data. 
numerical data being credit score, age, tenure, balance, products number, and estimated salary.
Categorical data being, country, gender, credit card (yes/no), active member (yes/no), and churn (yes/no)
we also observe that there is no missing data values or NA's

```{r}
summary(churned)

summary(notchurned)
```

the credit score range for not churned is smaller, yes, but the 1st Q, median, mean, 3rd Q, and max are too close for me to look hard at it.
the age max and min are similar, but the 4 inner numbers (1st Q, median, mean, 3rd Q) are vastly different, churned is much higher, this requires investigation
all 6 of the summary numbers in the tenure are very similar, no need.
I notice that the balance 1st Q for churned is 0, while the 1st Q for not churned is 38k, this could be a huge predictor.
There is a difference between the maxes of products owned, the pool is small with the mean being between 1 and 2 for both of them.
it is clear to see that most have a credit card, not much of a difference between those who did or did not churn.
the median for churned active member is 0 (no) and not churned active member is 1 (yes), the median for not churn being larger than churn too, this requires further investigation 

## Visuals

```{r}

#age
hist(churned$age, breaks = unique(churned$age), main = "Comparison of Churn and Not-Churn Ages", xlab = "Ages", col = rgb(0.8, 0.8, 0.8, 0.5), xlim = range(c(churned$age, notchurned$age)), freq = FALSE)
hist(notchurned$age, breaks = unique(notchurned$age), col = rgb(0.2, 0.2, 0.8, 0.5), add = TRUE, freq = FALSE) 
lines(density(churned$age), col = "red", lwd = 2)
lines(density(notchurned$age), col = "blue", lwd = 2)
legend("topright", 
       legend = c("Churned", "Not Churned"), 
       fill = c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)), 
       border = NA, 
       bty = "n")

#balance
boxplot(churned$balance, notchurned$balance, 
        horizontal = TRUE, 
        names = c("Churned", "Not Churned"), 
        col = c("red", "blue"),
        main = "Comparison of Balance Amounts", 
        xlab = "Balance Amounts")
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted")


#products number
PNcounts <- table(data$products_number)
barplot(PNcounts, main = "# of bank products", xlab = "amount", col = "grey")

#est salary
hist(churned$estimated_salary, 
     breaks = 50, 
     main = "Estimated Salaries Distribution (Churned vs Not Churned)", 
     xlab = "Salaries", 
     col = rgb(0.8, 0.8, 0.8, 0.5),
     xlim = range(c(churned$estimated_salary, notchurned$estimated_salary)), 
     freq = FALSE)
hist(notchurned$estimated_salary, 
     breaks = 50, 
     col = rgb(0.2, 0.2, 0.8, 0.5),
     add = TRUE, 
     freq = FALSE)
lines(density(churned$estimated_salary), col = "red", lwd = 2)
lines(density(notchurned$estimated_salary), col = "blue", lwd = 2)
legend("topright", 
       legend = c("Churned", "Not Churned"), 
       fill = c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)), 
       border = NA, 
       bty = "n") 


```
there is a obvious difference between churned and not churned ages and balance amounts
The large majority of the data has 1 or 2 bank products, I dont think its is reason enough to look at as a predictor
not so much of a difference between the estimated salaries, so I wont focus on it

```{r}
#country
churned_country_counts <- table(churned$country)
notchurned_country_counts <- table(notchurned$country)
country_counts_matrix <- rbind(churned_country_counts, notchurned_country_counts)
barplot(country_counts_matrix,
        beside = TRUE,
        col = c("red", "blue"), 
        main = "Country Distribution by Churn Status",
        xlab = "Country",
        ylab = "Count",
        legend.text = c("Churned", "Not Churned"), 
        args.legend = list(x = "topright")) 


# gender
churned_g_counts <- table(churned$gender)
notchurned_g_counts <- table(notchurned$gender)
g_counts_matrix <- rbind(churned_g_counts, notchurned_g_counts)
barplot(g_counts_matrix,
        beside = TRUE,
        col = c("red", "blue"), 
        main = "Gender Distribution by Churn",
        xlab = "Gender",
        ylab = "Count",
        legend.text = c("Churned", "Not Churned"), 
        args.legend = list(x = "top"))



#active member
churned_AM_counts <- table(churned$active_member)
notchurned_AM_counts <- table(notchurned$active_member)
AM_counts_matrix <- rbind(churned_AM_counts, notchurned_AM_counts)
barplot(AM_counts_matrix,
        beside = TRUE,
        col = c("red", "blue"), 
        main = "Active Member Distribution by Churn",
        xlab = "Active member",
        ylab = "Count",
        legend.text = c("Churned", "Not Churned"), 
        args.legend = list(x = "top"))

#churn
churncounts <- table(data$churn)
CHpercents <- round(100*churncounts/sum(churncounts), 1)
templabel <- c("no", "yes")
names(churncounts) <- templabel
CHlabels <-paste(templabel, churncounts, paste0( "(", CHpercents, "%)"))
pie(churncounts, labels = CHlabels, main = "Churn distribution ", col = c("grey", "black"))
```
for country churned/notchurn distribution, it looks like france and spain are about the same with 1/5th being churn and 4/5th not churned but for germany, 1/3rd is churned and 2/3rds is not churned, this could hint at something being up with germany.
for Gender, male and female look to be around the same with female churn count being slightly higher
the active member chart suggests that less active member are more likely to churn, which is intuitive.
for the data, 79.6% not churn. 20.4 yes churn


## stat tests

```{r}
# correlation test age vs balance
cor.test(data$balance, data$age)

# chi squared test
Contingency_table <- table(data$country, data$churn)
print(Contingency_table)
cst <- chisq.test(Contingency_table)
print(cst)

```

While the correlation is statistically significant (p-value < 0.05), it is not practically meaningful because the correlation coefficient is extremely close to 0.
Age and balance have almost no linear relationship. Factors other than age are likely more influential in determining a customer's balance.
this weak correlation suggests that age is not a strong predictor of balance.


the p value for country and churn is very small, which shows significance.
The significant p-value suggests there is a strong association between Country and Churn. Customers from different countries are not equally likely to churn.
Country could be a relevant predictor

the x^2 being 300+ shows that there is a big diference between the observed and expected frequencies.
