---
title: "model2_svm"
output:
  word_document: default
  html_document: default
date: "2024-12-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("caret")
library("kernlab")
library(ggplot2)

data <- read.csv("/Users/ahzenthu/Downloads/Bank Customer Churn Prediction.csv")
head(data)

data$country <- factor(data$country)
data$gender <- factor(data$gender)
data$churn <- factor(ifelse(as.numeric(data$churn) < 0.2, 'no', 'yes'))

sample_size <- floor(0.8 * nrow(data))
train_ind <- sample(seq_len(nrow(data)), size = sample_size)
train <- data[train_ind, ]
test <- data[-train_ind, ]

# had to change the CV bc my computer couldnt handle it well
fitControl <- trainControl(method = "cv", number = 3, verboseIter = TRUE)

grid <- expand.grid(degree = c(2), scale = c(0.01, 0.1), C = c(1))

svmpoly <- train(
  churn ~ credit_score + country + gender + age + tenure + balance + 
          products_number + credit_card + active_member + estimated_salary,
  data = train,
  method = "svmPoly",
  trControl = fitControl,
  tuneGrid = grid
)
print(svmpoly)

prediction_svmpoly<- predict(svmpoly, newdata = test)
confusionMatrix(prediction_svmpoly, test$churn)
```
the accuracy is high 85+, which show this model is pretty good.
the kappa shows moderate agreement, which shows that the model is strong.
the sensitivity is very high which shows its strong at predicting when customers will not churn
The model is excellent at identifying customers who will not churn (high sensitivity of 97).
It achieves a high overall accuracy of 86.05%.
The model struggles with identifying churn cases (low specificity of 36.6%).
The low specificity could lead to missed opportunities for targeting at-risk customers.

