---
title: "model1_LogR"
output:
  word_document: default
  html_document: default
date: "2024-12-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
data <- read.csv("/Users/ahzenthu/Downloads/Bank Customer Churn Prediction.csv")
head(data)

data$country <- factor(data$country)
data$gender <- factor(data$gender)

sample_size <- floor(0.8 * nrow(data))
train_ind <- sample(seq_len(nrow(data)), size = sample_size)
train <- data[train_ind, ]
test <- data[-train_ind, ]

glm.fit <- glm(
  churn ~ credit_score + country + gender + age + tenure + balance + 
           products_number + credit_card + active_member + estimated_salary,
  data = train,
  family = binomial
)

predictedprob <- predict(glm.fit, newdata = test, type = "response")

glm.pred <- ifelse(predictedprob > 0.5, "1", "0") 
glm.pred <- factor(glm.pred, levels = c("0", "1")) 

test$churn <- factor(test$churn, levels = c("0", "1"))
CM <- confusionMatrix(glm.pred, test$churn) # Predicted first, reference second
print(CM)
```
the model performs pretty well at predicting not churn cases with a high sensitivity and decent accuracy.
the model is weak at correctly predicting churn cases, seen from the low specificity.
the models precision is good at predicting not churn, 80+.
but when the model predicts churn, it is only right 60 percent of the time.
overall the model is better at predicting not churn than churn cases.

The model performs well for predicting non-churn cases due to high sensitivity (96.6%) and decent accuracy (82.1%).
It is good at predicting when a customer will not churn.

The model struggles to correctly predict churn cases, as seen from the low specificity (23.1%).
Low Kappa indicates the model is not very reliable when accounting for chance.

